---
title: "NHDFloline Subset"
author: "Michael Schmidt"
date: "February 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1 - Load Libraries 
If the libraries are not loaded on your computer you may have to run `install.packages("libraryName")`
```{r}
library(sf)
library(tidyverse)
```

## Step 1.2 - Assign variable values
```{r}
## No file extention for shapefiles to be read in
extent_file_path<-"inputs/shape/extent/little_snake_fo_extent"
flowlines_file_path<-"inputs/shape/flowlines/nhd_flowline_2018"

##where you want the function to put the files when they have been created. THIS FOLDER MUST ALREADY EXIST. 
output_file_path<-"inputs/shape/distance"
```

## Step 2 - Read in files and crop to field office extent
```{r}


flowlines<-st_read(dirname(flowlines_file_path), basename(flowlines_file_path))
crop_layer<-st_read(dirname(extent_file_path), basename(extent_file_path))
flowlines_sub<-st_crop(st_zm(flowlines), st_bbox(crop_layer))
```

## Step 3 - input a list with key value pairs of FCode number and related name for function below. 
```{r}
flowpath_type_list<-list(c(55800, "artificial"), c(46003, "intermittent"), c(46006, "perennial"))
```

## Step 4 - Add function to memory that subsets base on 
```{r}
stream_parser<-function(stream_id, stream_shape, output_path){
  stream_subset<-stream_shape%>%
    select(FCode)%>%
    filter(FCode==as.numeric(stream_id[1]))
  
  st_write(stream_subset, paste0(output_path,"/","nhd_flowline_", stream_id[2], ".shp"))
}
```


## Step 5 - Call function
```{r}
lapply(flowpath_type_list, stream_parser, flowlines_sub, output_file_path)
```

## Step 5 - Check to see if they are in the folder you specified.
